Project 1 for Complex Adaptive Systems

### 2018-02-20 paper.tex
Added piece on kernel map

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
The abstract goes here.
\end{abstract}

@@

\noindent For individual populations, we assume birth rate, death rate, and  carrying capacity are all fixed.  Define $k$ to be the carrying capacity of a population and $n_t$ to be size of the population at time $t$. Then the size of the population at generation $t$ is modeled by 
\begin{eqnarray}
n_{t+1} = (\textnormal{birthrate} - \textnormal{deathrate}) (k n_t - n_t^2) / k 
\end{eqnarray}
If we define the population size relative to the carrying capacity as  $x_t = n_t / k$ and define the reproductive fitness  $r = birthrate - deathrate$.  Then we  get the following equation for $x_t$: 
\begin{eqnarray}
x(t+1) = x_{t+1} = r  x_t (1 - x_t), \ \ \ \ \  t=0, 1, 2, \ldots . 
\label{logistic}
\end{eqnarray}
This is the logistic map as defined in \cite{Mitchell}.  In what follows, we will refer to $x_t$ as the population size, with the understanding that it will always be the ratio of the actual population size to the carrying capacity

@@

To compute the mutual information, we produced values that could be binned by multiplying each population size by 100 and rounding to the nearest integer.  Then we used the JIDT AutoAnalyzer \cite{JIDT} to compute the mutual information.    To generate figure \ref{figure1abcd} we used Matlab \cite{Matlab} with logistic map equation from \cite{Mitchell}.

MISSING {\it From instructions:  You might follow Flake?s suggestion of predicting how far into the future two populations are on the same side of 50\% of carrying capacity and use a binary representation. You might use JIDT?s kernel estimator. Explain your choice. }

@@

  \caption{ Population growth over time showing sensitive dependence on initial conditions.  The mutual information for the periodic map (a) is 0.7522.  The mutual information for the chaotic map (b) is 0.0471.}
 \label{figure1abcd}
\end{figure}

@@

\begin{figure*}[!t]
\centering
\subfloat[]{\includegraphics[width=3.5in]{Fig4a}%
\label{te}}
\hfill
\subfloat[]{\includegraphics[width=3.5in]{Fig4b}%
\label{fig_second_case}}
\caption{(a) Top-down, $T_{M\rightarrow X}$ (solid), and bottom-up, $T_{X\rightarrow M}$ (dashed), causal information transfer as a function of the global coupling parameter $\epsilon$ for coupled logistic maps. (b) The difference (Top-down minus bottom-up) in causal information transfer for $30$ coupled systems as well as the mean as a function of $\epsilon$ (dashed). Vertical line at $\epsilon = 0.3$ illustrates the directional reversal of causal information transfer.}
\label{te_avg}
\end{figure*}

@@

>>>>

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
We investigate information flow between higher and lower levels flow of information in models of population growth. 
 We attempt to replicate the results found in \cite{Walker} and DO OTHER STUFF. 
\end{abstract}

@@

\noindent For individual populations, we assume birth rate, death rate, and  carrying capacity are all fixed.  Define $k$ to be the carrying capacity of a population and $n_t$ to be size of the population at time $t$. Then the size of the population at generation $t$ is modeled by 
\begin{eqnarray}
n_{t+1} = (\textnormal{birthrate} - \textnormal{deathrate}) (k n_t - n_t^2) / k 
\end{eqnarray}
If we define the population size relative to the carrying capacity as  $x_t = n_t / k$ and define the reproductive fitness  $r = birthrate - deathrate$, then we  get the following equation for $x_t$: 
\begin{eqnarray}
x(t+1) = x_{t+1} = r  x_t (1 - x_t), \ \ \ \ \  t=0, 1, 2, \ldots . 
\label{logistic}
\end{eqnarray}
This is the logistic map as defined in \cite{Mitchell}.  In what follows, we will refer to $x_t$ as the population size, with the understanding that it will always be the ratio of the actual population size to the carrying capacity

@@

To estimate the mutual information, we produced values that could be binned by multiplying each population size by 100 and rounding to the nearest integer.  Then we used the JIDT AutoAnalyzer \cite{JIDT} to compute the mutual information assuming discrete data.  To generate figure \ref{figure1abcd} we used Matlab \cite{Matlab} with logistic map equation from \cite{Mitchell}.

@@

  \caption{ Population growth over time showing sensitive dependence on initial conditions.  The mutual information for the periodic map (a) is 0.7522.  The mutual information for the chaotic map (b) is 0.0471.}
 \label{figure1abcd}
\end{figure} 
To test whether binning the data might be adversely affecting our mutual information estimates and since our data was actually real instead of discrete, we also tried IDT?s kernel estimator to measure the mutual information. 
Kernel density estimators are a method for estimating probability densities \cite{Moon}. The JIDT kernel estimator assumes the input data is Gaussian and approximates distributions for all the variables based on this assumption.  One can then use the JIDT AutoAnalyzer to calculate the mutual information between the distributions.   Using the kernel estimator, we estimated the mutual information between the periodic maps to be 1.7086 and the mutual information between the chaotic maps to be 0.7192.   Again, this validates what we see in \ref{figure1abcd} because the chaotic maps have less mutual information than the periodic ones.

@@

\begin{figure*}[!t]
\centering
\subfloat[]{\includegraphics[width=3.5in]{Fig4a}%
\label{te} }

\hfill
\subfloat[]{\includegraphics[width=3.5in]{Fig4b}%
\label{fig_second_case} }
\caption{(a) Top-down, $T_{M\rightarrow X}$ (solid), and bottom-up, $T_{X\rightarrow M}$ (dashed), causal information transfer as a function of the global coupling parameter $\epsilon$ for coupled logistic maps. (b) The difference (Top-down minus bottom-up) in causal information transfer for $30$ coupled systems as well as the mean as a function of $\epsilon$ (dashed). Vertical line at $\epsilon = 0.3$ illustrates the directional reversal of causal information transfer.}
\label{te_avg}
\end{figure*}

@@

\bibitem{moon}
Moon, Young-Il, Balaji Rajagopalan, and Upmanu Lall.  Estimation of mutual information using kernel density estimators.  {\it Physical Review E}, Volume 52, Number 3, September 1995.
